{\fontfamily{ptm}\selectfont Linear Convergence of Stochastic Nesterov's Accelerated Proximal Gradient method under Interpolation Hypothesis, Truth or Just a Dream?}

### 2025-07-16 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, changse

\newcommand{\cov}{\ensuremath{\operatorname{Cov}}}
\newcommand{\var}{\ensuremath{\operatorname{Var}}}
\newcommand{\expect}{\ensuremath{\mathbb E}}
\usepackage{tikz}\newcommand*\circled[1]{\tikz[baseline=(char.base)]{\node[shape=circle,draw,inner sep=0.5pt](char) {#1};}}

@@

% ==============================================================================
\section{Introduction}
    \cite{bauschke_convex_2017}
    Previously we got some results, but unfortunately it was incorrect and it was impossible to recover from the mistake. 
    \par
    Does stochastic accelerated Nesterov's acceleration (SNAG) produces accelerated convergence rate (or, any type of convergence) when the Interpolation Hypothesis is true? 
    \textbf{I don't think that it's true after some mistakes from previous version of the notes and careful investigations.}
    In this file we develop some sufficient conditions for Linear convergence of (SNAG). 
    We will give explanations on why we don't think this is necessarily true. 
    \par
    When we use stochastic gradient to approximate the true graduate, it has an error. 
    Fix some $x\in \RR^n$, let $\tilde \nabla f(x)$ be an estimate of $\nabla f(x)$, the error we consider is $\expect \Vert \nabla f(x) - \nabla f(x)\Vert$. 
    To make the algebra simpler, we assume that the algorithm produced the next iterates $\tilde x$ by a step of gradient descent, and the error of the expectation satisfies a relative error conditions of the form 
    \begin{align*}
        \frac{
            \expect \left[
                \left\Vert \nabla f(x) - \tilde \nabla f(x) \right\Vert \Vert z - \tilde x\Vert
            \right]
            }{
            \expect \left[
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert
            \right]
            } &= \epsilon. 
    \end{align*}
    Where the variable $z$ will be explained later. 
    We will show that, the value of $\epsilon$ must decreases at a rate convergence relative to the Nesterov's accelerated sequence, under the standard Framework of analysis similar to what is in the literature. 
    Take note that usually in the literature, people analyze the quantity $\expect\left\Vert \tilde \nabla f(x) - \tilde \nabla f(y)\right\Vert^2$ for stochastic gradient type  of method. 
    The above expression is drastically different from what we usually have in the literature.

@@

            + \nabla f(x) + B(\tilde x - x).

@@

            Suppose that, a random variable $\tilde \nabla f(x)$ is used to estimate the proximal gradient operator and it produced an error $w$, and the estimate $\tilde x$. 

@@

                \right] &\le \epsilon \expect \left[

@@

                + \frac{\epsilon}{2} \left[\Vert x - \tilde x\Vert^2\right]. 

@@

@@

                \frac{B}{2}\Vert z - \tilde x\Vert &\le 

@@

@@

@@

@@

            And in this case, it's NOT dependended on $z$. 

@@

            \widetilde T_B^{(\epsilon)} = \argmin_{z \in \RR^n}\left\lbrace

@@

@@

        \textbf{Proof of (a)}. 

>>>>

\newcommand{\cov}{\ensuremath{\operatorname{Cov}}}
\newcommand{\var}{\ensuremath{\operatorname{Var}}}
\newcommand{\expect}{\ensuremath{\mathbb E}}
\usepackage{tikz}\newcommand*\circled[1]{
    \text{
        \hspace{-0.5em}\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=0.6pt](char){#1};
        }\hspace{-0.5em}
    }
}

@@

% ==============================================================================
\section{Introduction}
    \cite{bauschke_convex_2017}
    Previously we got some results, but unfortunately it was incorrect and, it was impossible to recover from the mistake. 
    \par
    Does stochastic accelerated Nesterov's acceleration (SNAG) produces accelerated convergence rate (or, any type of convergence) when the Interpolation Hypothesis is true? 
    \textbf{I don't think that it's true after some mistakes from previous version of the notes and careful investigations.}
    In this file we develop some sufficient conditions for Linear convergence of (SNAG). 
    We will give explanations on why we don't think this is necessarily true. 
    \par
    When we use stochastic gradient to approximate the true graduate, it has an error. 
    Fix some $x\in \RR^n$, let $\tilde \nabla f(x)$ be an estimate of $\nabla f(x)$, the error we consider is $\expect \Vert \nabla f(x) - \nabla f(x)\Vert$. 
    To make the algebra simpler, we assume that the algorithm produced the next iterates $\tilde x$ by a step of gradient descent, and the error of the expectation satisfies a relative error conditions of the form 
    \begin{align*}
        \frac{
            \expect \left[
                \left\Vert \nabla f(x) - \tilde \nabla f(x) \right\Vert \Vert z - \tilde x\Vert
            \right]
            }{
            \expect \left[
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert
            \right]
            } &= \epsilon. 
    \end{align*}
    Where the variable $z$ will be explained later. 
    We will show that, the value of $\epsilon$ must decreases at a rate convergence relative to the Nesterov's accelerated sequence, under the standard Framework of analysis similar to what is in the literature. 
    Take note that usually in the literature, people analyze the quantity $\expect\left\Vert \tilde \nabla f(x) - \tilde \nabla f(y)\right\Vert^2$ for stochastic gradient type  of method. 
    The above expression is drastically different from what we usually have in the literature.

@@

            + \nabla f(x) + B(\tilde x - x).\label{eqn:stoch-grad-err-vec}
        \end{align}

@@

            Suppose that, a random variable $\tilde \nabla f(x)$ is used to estimate the proximal gradient operator and, it produced an error $w$, and the estimate $\tilde x$ as defined in \ref{lemma:inex-pg-ineq-proto}. 

@@

                \right] &\le 
                \epsilon \expect \left[

@@

                + \frac{\epsilon}{2} \expect\left[\Vert x - \tilde x\Vert^2\right]. 

@@

            The reason for $w = \nabla f(x) - \tilde \nabla f(x)$ is explained in \eqref{eqn:stoch-grad-err-vec}. 

@@

@@

                F(z) - \expect F(\tilde x) + \frac{\max(B, L) - \mu}{2}\Vert z - x\Vert^2 

@@

                &\le F(z) - \expect F(\tilde x) + \frac{\max(B, L) - \mu}{2}\Vert z - x\Vert^2 
                + \epsilon \expect \left[

@@

                    \epsilon\Vert w\Vert\Vert z - \tilde x\Vert

@@

            And in this case, it doesn't on $z$. 

@@

            \widetilde T_B^{(\epsilon)}(x| F) = \argmin_{z \in \RR^n}\left\lbrace

@@

        For any fixed $\bar x \in \RR^n$. 
        Define $z_k = \alpha_k \bar x + (1 - \alpha_k)x_{k - 1}$ for all $k \ge 0$. 
        Consider the case for all $k \ge 1$. 
        Let $\expect_{k}$ be the expectation conditioned on all $\tilde \nabla f(y_i)$ for $i = 0, 1, \ldots, k - 1$. 
        We note that under this conditioning, the iterates $x_{k - 1}, v_{k - 1}, y_k$ are not random variables, but $x_k$, and $v_k$ are. 
        The sequence $(\alpha_k)_{k \ge 0}, (\epsilon_k)_{k \ge 0}$ are also not random variables. 
        We use Lemma \ref{lemma:stoch-pg-ineq} with $x = y_k, z = z_k, \tilde x = x_{k}$ and, $B = L$ then it has: 
            F(z_k) - \expect_{k - 1} F(x_k) 
            + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]
            &\underset{\circled{1}}{\le} 
            \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_{k - 1} F(x_k) - \frac{\mu\alpha_k(1 - \alpha_k)}{2}\Vert \bar x - x_{k - 1}\Vert^2
                + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]
            &\underset{\text{(d)}}{=}
            \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_{k - 1} F(x_k) 
                + \frac{(\alpha_k - 1)\mu(L\alpha_k - \mu)}{2(L - \mu)}\Vert x_{k - 1} - v_{k - 1}\Vert^2
                + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]
            \\&\underset{\textcolor{red}{(?)}}{\le}
            \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_{k - 1} F(x_k) 
                + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]
            \\&= 
            (1 - \alpha_k)(F(x_{k - 1}) - F(\bar x)) + F(\bar x) - \expect_{k - 1}F(x_k)
                + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]
            \\&= 
                F(x_{k - 1}) - F(\bar x)
                + \frac{\alpha_{k - 1}^2L\rho_{k - 1}}{2}\Vert \bar x - v_{k - 1}\Vert^2
                + F(\bar x) - \expect_{k - 1}F(x_k) 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]
            &\underset{\circled{2}}{=} 
                F(x_{k - 1}) - F(\bar x)
                + \frac{\alpha_{k - 1}^2L\rho_{k - 1}}{2}\Vert \bar x - v_{k - 1}\Vert^2
                + F(\bar x) - \expect_{k - 1}F(x_k) 
                - \frac{\alpha_k^2(L - \epsilon_k)}{2}\expect_{k - 1} \left[\Vert v_k - \bar x\Vert^2\right]. 
        The above produces the following inequality: 
        \begin{align}\begin{split}
            & \expect_{k - 1}F(x_k) - F(\bar x) - \frac{\alpha_k^2(L - \epsilon_k)}{2}\expect_{k - 1} \left[\Vert v_k - \bar x\Vert^2\right]
                F(x_{k - 1}) - F(\bar x)
                + \frac{\alpha_{k - 1}L\rho_{k - 1}}{2}\Vert \bar x - v_{k - 1}\Vert^2
            \right). 
        \end{split}\end{align}

@@

        \par \textbf{Proof of (a)}. 
